Do return series have power-law tails with the same index?

### 2017-04-30 papers/FX/draft.tex
Added to the proof of double sided pareto tail same tail parameters.
It appears the combined utility function is never monotonic.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where {\red where needed?}
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x \le  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
for the power-utility \fct\ \eqref{eq:u}.
A particularly interesting special case is when the left and the right
tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
K'$. In this case \eqref{eq:pareto} becomes
\begin{eqnarray*}
  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
  \int_{0}^\infty {
    u(C(x)) \left[
      1 - {b \over 1 + b}\1_{\{x > q\}}
      \right]
    + u(C(-x))
    \over
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)
\end{eqnarray*}
A detailed analysis of the montonicity of
\[
u(C(x)) \left[
  1 - {b \over 1 + b}\1_{\{x > q\}}
  \right]
+ u(C(-x))
\]
which comprises lemma \ref{lemma:III}, allows us to apply lemma
\ref{lemma:I} and conclude the following
\begin{itemize}
\item I am a good guy
\item So are you
\end{itemize}

@@

  1 - {b \over 1 + b}\1_{\{x > q\}}

@@

    \left(1 - {b \over 1 + b}\1_{\{x > q\}}\right)

@@

    \left( \1_{\{y \leq e^q\}} + c \1_{\{y > e^q\}} \right)

@@

      + \log(c)\1_{\{y > e^q\}} }_{g(y)} 

@@

    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > e^q\}}

@@

  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots

@@

      - 1)$ is positive.
      \item If $f(e^q) > g(e^q)$, then clearly $f(y) = g(y)$ has no
        solution on $(e^q, \infty)$; we also note $f(1) = g(1) =
        0$. Thus $f(y) = g(y)$ has no solution on $(1, e^q)$ either.

@@

        $(1, q)$ and $(q, \infty)$. To determine the monotonicity of

@@

      % conversely, $f(e^q) < g(e^q)$ implies $f(y) = g(y)$ must have a
      % solution on $(e^q, \infty)$, since \eqref{eq:f-g} tells us
      % $(f-g)(y) \to \infty$ as $y \to \infty$.

    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, $(f - g)(y)$
      is monotonically decreasing.
    \end{enumerate}

  % \item If $a < \min\{1, \xi\}$ or $a > \max\{1, \xi\})$, ${\pd (f
  %     -g)(y) \over \pd y} = 0$ has 2 real solutions.
  % \end{enumerate}

  % \begin{enumerate}
  % \item If $a < 1$ and $\xi > 1$
  %   \begin{enumerate}
  %   \item when $x > q$, $U'(x)$ being monotonic is equivalent to ${\pd U'(x)
  %       \over \pd x}$ having no zero on $(q, \infty)$.

  %     \begin{enumerate}
  %     \item If $f (e^q) > g(e^q)$ it is clear $f(y) = g(y)$
  %       has no solution on $(e^q, \infty)$.
  %     \item If $f (e^q) <  g(e^q)$ it is also clear $f(y) = g(y)$ must have
  %       one solution on $(e^q, \infty)$.
  %     \end{enumerate}
  %   \end{enumerate}

  %   \item If $a > 1$ and $\xi < 1$
  %     \begin{enumerate}
  %     \item when $x > q$, the same calculation that leads to
  %       \eqref{eq:fd} and \eqref{eq:gd} gives ${\pd f(y) \over \pd y}
  %       < 0$ and ${\pd g(y) \over \pd y} > 0$. So
  %     \begin{enumerate}
  %     \item If $f (e^q) >  g(e^q)$, $f(y) = g(y)$ must have
  %       one solution on $(e^q, \infty)$.
  %     \item If $f (e^q) < g(e^q)$, $f(y) = g(y)$ has no solution on $(e^q, \infty)$.
  %     \end{enumerate}
  %     \end{enumerate}

  %     \item If $a < 1$ and $\xi < 1$
  %       \begin{enumerate}
  %       \item when $x > q$
  %         \begin{enumerate}
  %         \item If $f (e^q) -  g(e^q) > 0$, we observe $f(y) - g(y)
  %           \to -\infty$ as $y \to \infty$. So by continuity $f(y) -
  %           g(y) = 0$ must have at least one solution on $(e^q, \infty)$.
  %         \end{enumerate}
  %       \end{enumerate}
  %     \item If $a > 1$ and $\xi > 1$
  %       \begin{enumerate}
  %       \item when $x > q$, both $f(y)$ and $g(y)$ decreases. Observe
  %         In this case, if $f(e^q) > g(e^q)$, there is no solution 
  %       \end{enumerate}
  \end{enumerate}
\end{proof}

>>>>

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where {\red where needed?}
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
for the power-utility \fct\ \eqref{eq:u}.
A particularly interesting special case is when the left and the right
tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
K'$. In this case \eqref{eq:pareto} becomes
\begin{eqnarray*}
  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
  \int_{0}^\infty {
    u(C(x)) \left[
      1 - {b \over 1 + b}\1_{\{x \ge q\}}
      \right]
    + u(C(-x))
    \over
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)
\end{eqnarray*}
A detailed analysis of the montonicity of
\[
u(C(x)) \left[
  1 - {b \over 1 + b}\1_{\{x \ge q\}}
  \right]
+ u(C(-x))
\]
which comprises lemma \ref{lemma:III}, allows us to apply lemma
\ref{lemma:I} and conclude the following
\begin{itemize}
\item I am a good guy
\item So are you
\end{itemize}

@@

  1 - {b \over 1 + b}\1_{\{x \ge q\}}

@@

    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)

@@

    \left( \1_{\{y < e^q\}} + c \1_{\{y > q\}} \right)

@@

      + \log(c)\1_{\{y > q\}} }_{g(y)} 

@@

    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > q\}}

@@

  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots when $\xi \ne 1$

@@

      - 1)$ is positive. We note $f(1) = g(1) = 0$; thus on $(1, e^q)$,
      $f(y) > g(y)$ and hence there is no solution to $f(y) = g(y)$ on
      this interval.
      \item If $ \lim_{y \to e^q}f(x) \ge \lim_{y \downarrow e^q} g(y)$,
        then clearly $f(y) = g(y)$ has no
        solution on $(e^q, \infty)$ either.

@@

        $(0, q)$ and $(q, \infty)$. To determine the monotonicity of

@@

        Thus ${\pd U'(x) \over \pd x} > 0$ for $x \in (0, q)$.
        To find out the sign of
        ${\pd U'(x) \over \pd x}$ on the interval $(q, \infty)$,
        one can look into the sign of ${\pd U'(x) \over \pd x}$
        as $x \to \infty$. For
        this purpose, substitute $x'$ for $e^{-x}$ and re-write
        ${\pd U'(x) \over \pd x}$ as
          {1 \over \xi}
          {\pd U'(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
          - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
        Taylor-expanding the last expression at $x'=0$ gives
          {1 \over \xi}
         = {x' \over a^{\xi + 1}} + o(x')
        From this expression we can see that
        ${\pd U'(x) \over \pd x} > 0$ for
        sufficiently large $x$; then by the proven monotonicity,
        ${\pd U'(x) \over \pd x} > 0$ for all $x \in (q, \infty)$.
        It remains to compare $\lim_{x\uparrow q}U'(x)$ with
        $\lim_{x\downarrow q}U'(x)$. Clearly
        \lim_{x\downarrow q}U'(x) - \lim_{x\uparrow q}U'(x) = - {
          (a + e^q)^{-\xi} b
          1+ b
        } < 0
        So $U'(x)$ and hence $U(x)$ as a whole is not monotonic on $(0, \infty)$.

      \item If $ \lim_{y \to e^q}f(x) < \lim_{y \downarrow e^q} g(y)$,
        $(f - g)(y)$ must have a solution on
        $(e^q, \infty)$ since $(f-g)(y)$ is monotonically
        increasing. So $U'(\cdot)$ is not monotonic in this case.
      \end{enumerate}

    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar argument as in the
      previous case, $(f - g)(y)$ is monotonically decreasing on both
      $(1, e^q)$ and $(e^q, \infty)$. From $f(1) = g(1) = 0$, it again
      follows that $U'(x)$ is monotonic on $(0, q)$.
      \begin{itemize}
      \item
        If $ \lim_{y \to e^q} f(y) > \lim_{y \downarrow e^q} g(y)$,
        then there must be a solution to $f(y) = g(y)$ on $(e^q,
        \infty)$. $U'(x)$ is not monotonic in this case.

      \item
        If $ \lim_{y \to e^q} f(y) \le \lim_{y \downarrow e^q} g(y)$,
        there is no solution to $f(y) = g(y)$ on $(e^q,
        \infty)$ either. $U'(x)$ is also monotonic on
        $(e^q, \infty)$. By the calculation as before, it can be seen
        ${\pd U'(x) \over \pd x} < 0$ for positive $x$ that is
        sufficiently close to 0; and
        ${\pd U'(x) \over \pd x} > 0$ for sufficiently large $x$.
        So $U'(x)$ is not monotonic in this case.
      \end{itemize}
    \end{enumerate}
  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
    is clear
    \[
    y_- = {
      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
      \over
      a (\xi - 1)
    } < 0
    \]
    It remains to compare $y_+$ with $1$ to determine whether $f(y) =
    g(y)$ has a solution. Assume $y_+ > 1$. Then
    \begin{eqnarray*}
      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1) \\
      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0
    \end{eqnarray*}
    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
    on both $(1, e^q)$ and $(e^q, \infty)$. Following the same
    analysis as in the case (1.a.i), one can see $U'(\cdot)$ is not
    monotonic on $(0, \infty)$.

  \item If $a < \min\{1, \xi\}$ and $\xi < 1$, i.e. $a < \xi < 1$, it
    is clear $y_- > 0$ and $y_+ < y_-$. Moreover, $y_- > 1$ is equivalent
    to
    \[
    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi = (a + 1) (a - \xi)
    \]
    The last inequality is obviously true in this case. So $y_- >
    1$. ${\pd (f-g)(y) \over \pd y}$ has a maximum at
    $y_- \in (1, \infty)$. We know $(f - g)(y_-)$ is a maximum because
    \eqref{eq:f-g-der} shows that, for $y > y_-$,
    ${\pd (f-g)(y) \over \pd y} < 0$.

    If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
    \infty)$. Hence $U'(\cdot)$ is monotonic on both $(0, q)$ and $(q,
    \infty)$. The same analysis as in the case (1.a.i) shows $U'(\cdot)$
    is not monotonic on $(0, \infty)$.

    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
    \infty)$. So $U'(\cdot)$ is not monotonic.
  \end{enumerate}
\end{proof}
